Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-04-09 src/tex/ms.tex
[showyourwork] overleaf sync

The power-law ($\alpha=-1.1$) DTD has the strongest observational motivation but poorly reproduces the disk abundance distributions. This can be mitigated somewhat with a longer minimum delay time, which has a similar effect on chemical evolution tracks as the addition of an initial plateau in the DTD (see discussion in Section \ref{sec:onezone-results}). Even so, it is clear that the high fraction of prompt SNe Ia in the DTD of, e.g., \citet{MaozMannucci2012-SNeIaReview} is at odds with Galactic chemical abundance measurements. 
Measurements of the cosmic SN Ia rate become considerably uncertain at $z\gtrsim1$ \citep[see, e.g.,][]{Palicio2024-CosmicSNIaRate}. Additionally, measurements of galactic or cosmic SFHs typically provide constraints for the DTD in coarse age bins, with especially large uncertainties in the youngest bins \citep[e.g.,][]{MaozMannucci2012-SNeIaReview},
%, and it is difficult to constrain the SFH of individual galaxies below [number] [cite],
so constraints on the DTD from external galaxies should be more sensitive to the rates at long delay times. On the other hand, our results demonstrate that the high-$\alpha$ sequence in GCE models is highly sensitive to the DTD at short delay times.

@@

% Is the MW DTD the same as elsewhere?
Our finding that the $t^{-1}$ power-law DTD poorly reproduces observed abundances is a signal that the Milky Way could obey a different DTD than other galaxies. This would not be too far beyond \citet{Maoz2017-CosmicDTD}'s finding that field galaxies and galaxy clusters have a different DTD slope. On the other hand, \citet{Walcher2016-SelfSimilarity} argue that the similarity of the age--\aFe relation between solar neighborhood stars and nearby elliptical galaxies is evidence for a universal DTD. A physical mechanism would be needed to produce a different slope or form for the DTD in different environments, such as a metallicity dependence in the fraction of close binaries \citep[e.g.,][]{Moe2019-CloseBinaryFraction}.

>>>>

The power-law ($\alpha=-1.1$) DTD has the strongest observational motivation but poorly reproduces the disk abundance distributions. This can be mitigated somewhat with a longer minimum delay time, which has a similar effect on chemical evolution tracks as the addition of an initial plateau in the DTD (see discussion in Section \ref{sec:onezone-results}). Even so, it is clear that the high fraction of prompt SNe Ia in the DTD of, e.g., \citet{MaozMannucci2012-SNeIaReview} is at odds with Galactic chemical abundance measurements.

Measurements of the cosmic SN Ia rate become considerably uncertain at $z\gtrsim1$ \citep[see, e.g.,][]{Palicio2024-CosmicSNIaRate}. Additionally, measurements of galactic or cosmic SFHs typically provide constraints for the DTD in coarse age bins, with especially large uncertainties in the youngest bins \citep[e.g.,][]{MaozMannucci2012-SNeIaReview},
%, and it is difficult to constrain the SFH of individual galaxies below [number] [cite],
so constraints on the DTD from external galaxies should be more sensitive to the rates at long delay times. On the other hand, our results demonstrate that the high-$\alpha$ sequence in GCE models is highly sensitive to the DTD at short delay times.

@@

% Is the MW DTD the same as elsewhere?
Our finding that the $t^{-1}$ power-law DTD poorly reproduces observed abundances could suggest that the Milky Way obeys a different DTD than other galaxies. This would not be too far beyond \citet{Maoz2017-CosmicDTD}'s finding that field galaxies and galaxy clusters have a different DTD slope. On the other hand, \citet{Walcher2016-SelfSimilarity} argue that the similarity of the age--\aFe relation between solar neighborhood stars and nearby elliptical galaxies is evidence for a universal DTD. A physical mechanism would be needed to produce a different slope or form for the DTD in different environments, such as a metallicity dependence in the fraction of close binaries \citep[e.g.,][]{Moe2019-CloseBinaryFraction}.
